---
title: "STA303H1 Final Project"
author: "Danish Ahmed"
date: "2024-04-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Final Project Part 2
##### Setting up the document and reading the dataset
```{r}
library(dplyr)
library(ggplot2)
library(rms)
df <- read.csv('train.csv')
head(df)
```

##### Data Cleaning and changing data types for further analysis
```{r}
# removing all the observations with null values
df <- na.omit(df)
df$owns_car <- ifelse(df$owns_car == "Y", 1, 0)
df$owns_house <- ifelse(df$owns_house == "Y", 1, 0)
df$gender <- ifelse(df$gender == "M", 1, 0)
head(df)
```
### EDA Figures
##### Credit Default based on gender
```{r}
# Assuming 0 represents male and 1 represents female in the gender column
# Assuming 1 represents Yes and 0 represents No in the credit_default column

# Create a new column for gender labels based on the gender column values
df$gender_label <- ifelse(df$gender == 1, "Male", "Female")

# Create the bar plot with dodged position
ggplot(df, aes(x = gender_label, fill = factor(credit_card_default))) +
  geom_bar(position = "dodge") +
  labs(title = "Credit Card Default by Gender", x = "Gender", y = "Number of Defaults", fill = "Credit Card Default") +
  scale_fill_manual(values = c("lightblue", "lightcoral"), labels = c("No", "Yes"))

```

##### Credit Default bassed on Migrant Worker Status
```{r}
library(ggplot2)

# Assuming 1 represents migrant workers and 0 represents non-migrant workers in the migrant_worker column
# Assuming 1 represents Yes and 0 represents No in the credit_default column

# Create a new column for worker labels based on the migrant_worker column values
df$worker_label <- ifelse(df$migrant_worker == 1, "Migrant Worker", "Non-Migrant Worker")

# Create the bar plot with dodged position
ggplot(df, aes(x = worker_label, fill = factor(credit_card_default))) +
  geom_bar(position = "dodge") +
  labs(title = "Credit Card Default by Worker Type", x = "Worker Type", y = "Number of Defaults", fill = "Credit Card Default") +
  scale_fill_manual(values = c("lightblue", "lightcoral"), labels = c("No", "Yes"))

```

```{r}
# Histogram for credit_limit_used
ggplot(df, aes(x = credit_limit_used...)) +
  geom_histogram(binwidth = 10, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Credit Limit Used(%)", x = "Credit Limit Used(%)", y = "Number of People")
```
```{r}
# Histogram for Number of days employed
df <- df %>%
  filter(no_of_days_employed < 10000)
ggplot(df, aes(x = no_of_days_employed)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of the Total Number of Days Employed", x = "Days Employed", y = "Number of People")
```

##### Initial Model Check with all variables
```{r}
# Regular model check
logit.mod1 <- glm(credit_card_default ~ age + gender + owns_car + owns_house + no_of_children + net_yearly_income + no_of_days_employed + total_family_members + migrant_worker + yearly_debt_payments + credit_limit + credit_limit_used... + credit_score + prev_defaults + default_in_last_6months + gender * owns_car + credit_limit * net_yearly_income + gender * owns_house + migrant_worker * no_of_days_employed, family = binomial(link = logit), data = df)
summary(logit.mod1)
```

# Stepwise elimination based on AIC to choose variables
```{r}
# There are no missing values in the data and so we do not have to remove any observations
#install.packages('sjPlot')

logit.mod1 <- glm(credit_card_default ~ age + gender + owns_car + owns_house + no_of_children + net_yearly_income + no_of_days_employed + total_family_members + migrant_worker + yearly_debt_payments + credit_limit + credit_limit_used... + credit_score + prev_defaults + default_in_last_6months + gender * owns_car + credit_limit * net_yearly_income + gender * owns_house + migrant_worker * no_of_days_employed, family = binomial(link = logit), data = df)

sel.var.aic <- step(logit.mod1, trace = 0, k = 2, direction = "both") 
select_var_aic<-attr(terms(sel.var.aic), "term.labels")   
select_var_aic
```


# Stepwise elimination based on BIC to choose variables
```{r}
sel.var.bic <- step(logit.mod1, trace = 0, k = log(nrow(df)), direction = "both") 
select_var_bic<-attr(terms(sel.var.bic), "term.labels")   
select_var_bic
```

# LASSO elimination to choose variables


```{r}
# Load necessary libraries
library(glmnet)
x <- model.matrix(credit_card_default ~ age + gender + owns_car + owns_house + no_of_children + net_yearly_income + no_of_days_employed + total_family_members + migrant_worker + yearly_debt_payments + credit_limit + credit_limit_used... + credit_score + prev_defaults + default_in_last_6months + gender * owns_car + credit_limit * net_yearly_income + gender * owns_house + migrant_worker * no_of_days_employed -1, family = binomial(), 
    data = df) # -1 to exclude the intercept
y <- df$credit_card_default
cv.out = cv.glmnet(x, y, family = "binomial", type.measure = "class", alpha = 1)

plot(cv.out)
best.lambda <- cv.out$lambda.1se
best.lambda
co<-coef(cv.out, s = "lambda.1se")

#Selection of the significant features(predictors)

## threshold for variable selection ##

thresh <- 0.00
# select variables #
inds<-which(abs(co) > thresh )
variables<-row.names(co)[inds]
sel.var.lasso<-variables[!(variables %in% '(Intercept)')]
sel.var.lasso
```

##### Building the first model according to Stepwise AIC variable selection with the lowest AIC as the optimum model as output
```{r}
logit.mod.AIC <- glm(credit_card_default ~ gender + owns_car + owns_house + no_of_children + net_yearly_income + no_of_days_employed + total_family_members + migrant_worker + credit_limit + credit_limit_used... + credit_score + prev_defaults + net_yearly_income:credit_limit + gender:owns_house + no_of_days_employed:migrant_worker, family = binomial(link = logit), data = df)
summary(logit.mod.AIC)
```

# Building the model according to Stepwise BIC variable selection
```{r}
logit.mod.BIC <- glm(credit_card_default ~ gender + owns_car + no_of_days_employed + 
    migrant_worker + credit_limit_used... + credit_score + prev_defaults + 
    no_of_days_employed:migrant_worker, family = binomial(link = logit), data = df)
summary(logit.mod.BIC)
```

# Building the model according to LASSO variable selection
```{r}
logit.mod.LASSO <- glm(credit_card_default ~ credit_limit_used... + credit_score + prev_defaults + default_in_last_6months, family = binomial(link = logit), data = df)
summary(logit.mod.LASSO)
```

# Cross-validation and calibration plot for model generated through Stepwise AIC variable selection
```{r}
library(rms)
logit.mod.AIC.cross <- lrm(credit_card_default ~ age + gender + owns_car + owns_house + no_of_children + net_yearly_income + no_of_days_employed + total_family_members + migrant_worker + credit_limit + credit_limit_used... + prev_defaults + gender * owns_house + no_of_days_employed * migrant_worker, data = df, x=TRUE, y=TRUE, model = T, maxit=1000)
```

```{r}
cross.calib <- calibrate(logit.mod.AIC.cross, method="crossvalidation", B=10) # model calibration
plot(cross.calib, las=1, xlab = "Predicted Probability")
```


# Cross-validation and calibration plot for model generated through Stepwise BIC variable selection
```{r}
logit.mod.BIC.cross <- lrm(credit_card_default ~ gender + owns_car + no_of_days_employed + migrant_worker + credit_limit_used... + prev_defaults + migrant_worker * no_of_days_employed, data = df, x=TRUE, y=TRUE, model = T)
```

```{r}
cross.calib <- calibrate(logit.mod.BIC.cross, method="crossvalidation", B=10) # model calibration
plot(cross.calib, las=1, xlab = "Predicted Probability")
```

# Cross-validation and calibration plot for model generated through LASSO variable selection

```{r}
logit.mod.LASSO.cross <- lrm(credit_card_default ~ gender + owns_car + no_of_days_employed + migrant_worker + credit_limit_used... + prev_defaults + default_in_last_6months + no_of_days_employed:migrant_worker, data = df, x=TRUE, y=TRUE, model = T)
cross.calib <- calibrate(logit.mod.LASSO.cross, method="crossvalidation", B=10) # model calibration
plot(cross.calib, las=1, xlab = "Predicted Probability")
```


# Producing ROC for model generated by Stepwise AIC

```{r}
#install.packages('pROC')
library(pROC)
p <- predict(logit.mod.AIC.cross, type = "fitted")

roc_logit <- roc(df$credit_card_default ~ p)
## The True Positive Rate ##
TPR <- roc_logit$sensitivities
## The False Positive Rate ##
FPR <- 1 - roc_logit$specificities

plot(FPR, TPR, xlim = c(0,1), ylim = c(0,1), type = 'l', lty = 1, lwd = 2,col = 'red')
abline(a = 0, b = 1, lty = 2, col = 'blue')
text(0.7,0.4,label = paste("AUC = ", round(auc(roc_logit),2)))

auc(roc_logit)
```


# Producing ROC for model generated by Stepwise BIC
```{r}
library(pROC)
p <- predict(logit.mod.BIC.cross, type = "fitted")

roc_logit <- roc(df$credit_card_default ~ p)
## The True Positive Rate ##
TPR <- roc_logit$sensitivities
## The False Positive Rate ##
FPR <- 1 - roc_logit$specificities

plot(FPR, TPR, xlim = c(0,1), ylim = c(0,1), type = 'l', lty = 1, lwd = 2,col = 'red')
abline(a = 0, b = 1, lty = 2, col = 'blue')
text(0.7,0.4,label = paste("AUC = ", round(auc(roc_logit),2)))

auc(roc_logit)
```

# Producing ROC for model generated by LASSO

```{r}
library(pROC)
p <- predict(logit.mod.LASSO.cross, type = "fitted")

roc_logit <- roc(df$credit_card_default ~ p)
## The True Positive Rate ##
TPR <- roc_logit$sensitivities
## The False Positive Rate ##
FPR <- 1 - roc_logit$specificities

plot(FPR, TPR, xlim = c(0,1), ylim = c(0,1), type = 'l', lty = 1, lwd = 2,col = 'red')
abline(a = 0, b = 1, lty = 2, col = 'blue')
text(0.7,0.4,label = paste("AUC = ", round(auc(roc_logit),2)))

auc(roc_logit)
```


# DFFITS for model generated through BIC variable selection as final model is BIC taken
```{r}

dffits_values <- dffits(logit.mod.BIC)

eta <- predict(logit.mod.BIC, type = "link")
par(family = 'serif')

plot(eta, dffits_values, xlab='Linear Predictor', 
     ylab='DFFITS', pch=20)

lines(lowess(eta, dffits_values), lwd=2, col='blue')

abline(h=0, lty='dotted')

threshold <- 2 * sqrt(length(coef(logit.mod.LASSO))/nrow(df))
abline(h=-threshold, lty='dotted')
abline(h=threshold, lty='dotted')
```


